---
title: "Introduction to R/ggplot"
output:
  slidy_presentation:
    font_adjustment: -1
---

Getting Started
===============

* Run `R` for interactive REPL

* Getting help

    ```{r, eval=FALSE}
?length
?'+'
??zip
```

* Installing libraries

    ```{r, eval=FALSE}
install.packages('ggplot2')
```

* Run scripts using `Rscript` from command line

    ```sh
Rscript plots.R
```


R Data Types
============
* primitives
* vectors
* matrices (won't cover)
* lists
* data frames

Primitives
==========
```{r}
3/2
TRUE
T
FALSE
F
'hello'
sprintf('I have %d apples and %0.02f bananas', 2, pi)
```

Vectors
=======
```{r}
(vector = c(5, 3, 9)) # extra parens to print value
length(vector)
```

Flat, uniform:
```{r}
c(5, 3, 9, FALSE, c(7, 2))
10:20
rep(7, 5)
seq(10, 12, .5)
```

Nearly all operations vectorized:
```{r}
vector * 3
vector + 6
vector == 3
```

Lists
=====

* Sometimes a (Python) list, sometimes a map
* Like a PHP array

```{r}
values = list(1, 'apple', 4, 'me')
values[[2]]
mapping = list(x=3, y=TRUE, values=values)
mapping[['x']]
mapping$y
mapping$values[[4]]
ls(mapping)
```

Data Frames
===========

* Like a database table

    ```{r, results='asis'}
df = data.frame(x=c('a', 'b'), y=c(1, 2))
fmt = knitr::kable # renders data frame nicely
fmt(df)
df$y

somecars = mtcars[1:5,]
fmt(somecars)
```

*  Projection - vector of column names after the comma

    ```{r, results='asis'}
fmt(somecars[, c('mpg', 'hp', 'gear')]) # projection
```

* Selection - vector of booleans before the comma

    ```{r, results='asis'}
fmt(somecars[somecars$mpg > 21, ])
```

* Combined

    ```{r, results='asis'}
fmt(somecars[somecars$mpg > 21, c('hp', 'gear')])
```
